---
title: "Truth, beauty, and data + R and tidyverse"
date: "2022-06-06"
date_end: "2022-06-10"
---

```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))
source(here::here("R", "youtube-playlist.R"))

playlist_id <- "PLS6tnpTr39sHom5NPjrZAeMTYS5orTl9c"

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "tVcomh6jQ2Y", "Introduction", "1", TRUE,
  "UbQ8IW3UI9E", "Facts, truth, and beauty", "facts-truth-beauty", FALSE,
  "DldFVc08YY4", "Data, truth, and beauty", "data-truth-beauty", FALSE,
  "6MTC9uGMNFg", "Beautiful visualizations", "beautiful-visualizations", FALSE,
  "_TclSs8iLMY", "Class details", "class-details", FALSE
)
```

## Readings

- The [syllabus](/syllabus/), [content](/content/), [lessons](/lesson/), [examples](/example/), and [assignments](/assigment/) pages for this class
- {{< fa podcast >}} Tim Harford, ["Florence Nightingale: Data Viz Pioneer,"](https://99percentinvisible.org/episode/florence-nightingale-data-viz-pioneer/) *99% Invisible*, episode 433, March 2, 2021 (*this is a podcast; listen to it in your browser or use an app like [Overcast](https://overcast.fm/) or [Spotify](https://spotify.com/)*)
- {{< fa book >}} [Chapter 1](http://socviz.co/lookatdata.html) in *Data Visualization* [@Healy:2018]
- {{< fa book >}} Chapters 2 and 3 in *The Truthful Art* [@Cairo:2016] (*skim the introduction and chapter 1*)
- {{< fa arrow-up-right-from-square >}} [Study: Charts change hearts and minds better than words do](https://www.washingtonpost.com/news/wonk/wp/2018/06/15/study-charts-change-hearts-and-minds-better-than-words-do/?utm_term=.4474599c0d5e)


### Questions to reflect on

*(Remember, you don't need to answer all of these—or even any of them! These are just here to help guide your thinking.)*

- How do we know what is true?
- Are facts truth?
- Why do we visualize data?
- What makes a great visualization?
- How do you choose which kind of visualization to use?


## Slides

The slides for today's lesson are available online as an HTML file. Use the buttons below to open the slides either as an interactive website or as a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/01-slides")

xaringanExtra::embed_xaringan(
  "/slides/01-slides.html",
  ratio = "16:9",
  border = "none",
  margin = "0 auto"
)
```

:::{.callout-tip}
**Fun fact**: If you type <kbd>?</kbd> (or <kbd>shift</kbd> + <kbd>/</kbd>) while going through the slides, you can see a list of special slide-specific commands.
:::


## Videos

```{r show-youtube-list, echo=FALSE, results="asis"}
youtube_list(slide_details, playlist_id)
```
