---
title: "Graphic design"
date: "2022-06-06"
date_end: "2022-06-10"
---

```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))
source(here::here("R", "youtube-playlist.R"))

playlist_id <- "PLS6tnpTr39sEznTwka0EmWfkkphjncq7U"

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "ny7HYqyVNp4", "Introduction", "1", TRUE,
  "SXW0RtenLgk", "Truth, beauty, stories, design", "truth-beauty-design", FALSE,
  "iuEwh2EnIcw", "Graphic design and CRAP", "design-crap", FALSE,
  "j0aqBmRV66A", "Image types", "image-types", FALSE
)
```

## Readings

- {{< fa book >}} Chapter 5 in Alberto Cairo, *The Truthful Art* [@Cairo:2016]
- {{< fa book >}} [Chapter 4](https://clauswilke.com/dataviz/color-basics.html) and [Chapter 27](https://clauswilke.com/dataviz/image-file-formats.html) in Claus Wilke, *Fundamentals of Data Visualization* [@Wilke:2018]
- {{< fa arrow-up-right-from-square >}} Cédric Scherer, ["Colors and emotions in data visualization"](https://blog.datawrapper.de/colors-for-data-vis-style-guides/)
- {{< fa arrow-up-right-from-square >}} Lisa Charlotte Muth, ["A detailed guide to colors in data vis style guides"](https://www.storytellingwithdata.com/blog/2021/6/8/colors-and-emotions-in-data-visualization). You can skim this—this is a good resource for future you.
- {{< fa file-pdf >}} [Summary of CRAP graphic design principles](http://www.presentationzen.com/chapter6_spread.pdf) from Garr Reynolds, *Presentation Zen* [@Reynolds:2008]. These principles are from Robin Williams' [*The Non-Designer's Design & Type Books*](https://www.amazon.com/Non-Designers-Design-Book-4th/dp/0133966151) [@Williams:2008], which you should really get if you're interested in doing anything design-related ever. Her stuff is life-changing.
- {{< fa arrow-up-right-from-square >}} [Typography in ten minutes](https://practicaltypography.com/typography-in-ten-minutes.html). The rest of the *Practical Typography* book is phenomenal and you'd be remiss if you didn't read the whole thing and bookmark it for life, but for now just read this quick summary.
- {{< fa arrow-up-right-from-square >}} ["What's the Difference Between JPG, PNG, and GIF?"](https://gizmodo.com/5656669/whats-the-difference-between-jpg-png-and-gif)
- {{< fa arrow-up-right-from-square >}} ["File formats explained"](https://www.theglowstudio.com/file-formats-explained/)

### Questions to reflect on

*(Remember, you don't need to answer all of these—or even any of them! These are just here to help guide your thinking.)*

- Why does graphic design matter when conveying truth?
- What makes something well designed (vs. poorly designed)?


## Slides

The slides for today's lesson are available online as an HTML file. Use the buttons below to open the slides either as an interactive website or as a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/02-slides")
slide_tabs(slide_details, "/slides/02-slides.html")
```

:::{.callout-tip}
**Fun fact**: If you type <kbd>?</kbd> (or <kbd>shift</kbd> + <kbd>/</kbd>) while going through the slides, you can see a list of special slide-specific commands.
:::


## Videos

```{r show-youtube-list, echo=FALSE, results="asis"}
youtube_list(slide_details, playlist_id)
```
