---
title: "Uncertainty"
date: "2022-06-20"
date_end: "2022-06-24"
---

```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))
source(here::here("R", "youtube-playlist.R"))

playlist_id <- "PLS6tnpTr39sFmy596lUElQakCrKDSHUUU"

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "eZTp8Hninv4", "Introduction", "1", TRUE,
  "TCQwkwlj4kM", "Communicating uncertainty", "communicating", FALSE,
  "X6zPEOPo-2M", "Visualizing uncertainty", "visualizing", FALSE
)
```

## Readings

- {{< fa book >}} [Chapter 7](https://clauswilke.com/dataviz/histograms-density-plots.html) in Claus Wilke, *Fundamentals of Data Visualization* [@Wilke:2018]
- {{< fa book >}} Chapters 4, 7, and 11 in Alberto Cairo, *The Truthful Art* [@Cairo:2016]
- {{< fa arrow-up-right-from-square >}} [Why It's So Hard for Us to Visualize Uncertainty](https://hbr.org/2016/11/why-its-so-hard-for-us-to-visualize-uncertainty)
- {{< fa brands youtube >}} [Amanda Cox's keynote address at the 2017 OpenVis Conf](https://www.youtube.com/watch?v=0L1tGo-DvD0)
- {{< fa arrow-up-right-from-square >}} [Communicating Uncertainty When Lives Are on the Line](https://eagereyes.org/blog/2017/communicating-uncertainty-when-lives-are-on-the-line)
- {{< fa arrow-up-right-from-square >}} [Showing uncertainty during the live election forecast](https://flowingdata.com/2016/11/15/showing-uncertainty-during-the-live-election-forecast/) & [Trolling the uncertainty dial](https://flowingdata.com/2017/06/27/trolling-the-uncertainty-dial/)
- {{< fa arrow-up-right-from-square >}} Cédric Scherer, ["Visualizing distributions with raincloud plots with ggplot2"](https://www.cedricscherer.com/2021/06/06/visualizing-distributions-with-raincloud-plots-with-ggplot2/)

### Questions to reflect on

*(Remember, you don't need to answer all of these—or even any of them! These are just here to help guide your thinking.)*

- Why is it important to deal with uncertainty in data?
- What was good or bad about the New York Times' 2016 live election guage?
- Why is it so hard to visualize uncertainty?
- Why is it so hard to communicate uncertainty to others?


## Slides

The slides for today's lesson are available online as an HTML file. Use the buttons below to open the slides either as an interactive website or as a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/06-slides")
slide_tabs(slide_details, "/slides/06-slides.html")
```

:::{.callout-tip}
**Fun fact**: If you type <kbd>?</kbd> (or <kbd>shift</kbd> + <kbd>/</kbd>) while going through the slides, you can see a list of special slide-specific commands.
:::


## Videos

```{r show-youtube-list, echo=FALSE, results="asis"}
youtube_list(slide_details, playlist_id)
```


```{cat include=FALSE, eval=FALSE}
https://www.youtube.com/playlist?list=PLS6tnpTr39sFmy596lUElQakCrKDSHUUU

https://youtu.be/eZTp8Hninv4
PMAP 8921 • (6) Uncertainty: (0) Introduction

Introduction to the class session on uncertainty.

Access the slides, code, and other materials at the course website at https://datavizm20.classes.andrewheiss.com/content/06-content/


https://youtu.be/TCQwkwlj4kM
PMAP 8921 • (6) Uncertainty: (1) Communicating uncertainty

Discussion about why uncertainty is both important and difficult to communicate 

Access the slides, code, and other materials at the course website at https://datavizm20.classes.andrewheiss.com/content/06-content/


https://youtu.be/X6zPEOPo-2M
PMAP 8921 • (6) Uncertainty: (2) Visualizing uncertainty

Overview of different methods for visualizing uncertainty, including histograms, density plots, ridge plots, box plots, violin plots, and half-versions of all these plots

Access the slides, code, and other materials at the course website at https://datavizm20.classes.andrewheiss.com/content/06-content/


https://youtu.be/40dW63jTbsk
PMAP 8921 • (6) Uncertainty: Example

Demonstration of how to create histograms, density plots, ridge plots, box plots, violin plots with ggplot2 and R

Download the data and see the polished code at https://datavizm20.classes.andrewheiss.com/example/06-example/
  
```
