---
title: "Annotations"
date: "2022-07-05"
date_end: "2022-07-08"
---

```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))
source(here::here("R", "youtube-playlist.R"))

playlist_id <- "PLS6tnpTr39sG4NCNqLqHhIE7khPgrdRJD"

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "1c51k2tmjBI", "Introduction", "1", TRUE,
  "6p4K5_6KU9k", "Fretting the little things", "little-things", FALSE,
  "fdR78oXb_fI", "Text in plots", "text", FALSE,
  "lV_Qak75E6I", "Seeds", "seeds", FALSE
)
```

## Readings

- {{< fa book >}} [Chapter 17](https://clauswilke.com/dataviz/redundant-coding.html), [Chapter 18](https://clauswilke.com/dataviz/multi-panel-figures.html), [Chapter 19](https://clauswilke.com/dataviz/figure-titles-captions.html), and [Chapter 21](https://clauswilke.com/dataviz/small-axis-labels.html) in *Fundamentals of Data Visualization* [@Wilke:2018]
- {{< fa book >}} [Chapter 5](http://socviz.co/workgeoms.html) in *Data Visualization* [@Healy:2018]
- {{< fa arrow-up-right-from-square >}} Cara Thompson, ["Level Up Your Labels: Tips and Tricks for Annotating Plots"](https://www.cararthompson.com/talks/user2022)


## Slides

The slides for today's lesson are available online as an HTML file. Use the buttons below to open the slides either as an interactive website or as a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/09-slides")

xaringanExtra::embed_xaringan(
  "/slides/09-slides.html",
  ratio = "16:9",
  border = "none",
  margin = "0 auto"
)
```

:::{.callout-tip}
**Fun fact**: If you type <kbd>?</kbd> (or <kbd>shift</kbd> + <kbd>/</kbd>) while going through the slides, you can see a list of special slide-specific commands.
:::


## Videos

```{r show-youtube-list, echo=FALSE, results="asis"}
youtube_list(slide_details, playlist_id)
```


```{cat include=FALSE, eval=FALSE}
https://www.youtube.com/playlist?list=PLS6tnpTr39sG4NCNqLqHhIE7khPgrdRJD

https://youtu.be/1c51k2tmjBI
PMAP 8921 • (9) Annotations: (0) Introduction

Introduction to the class session on annotations

Access the slides, code, and other materials at the course website at https://datavizm20.classes.andrewheiss.com/content/09-content/


https://youtu.be/6p4K5_6KU9k
PMAP 8921 • (9) Annotations: (1) Fretting the little things

Discussion of why it's important to fret about little details when designing plots and how little tweaks can enhance the accessibility and usability of your plots

Access the slides, code, and other materials at the course website at https://datavizm20.classes.andrewheiss.com/content/09-content/


https://youtu.be/fdR78oXb_fI
PMAP 8921 • (9) Annotations: (2) Text in plots

Overview and examples of three main ways of adding text to ggplot plots in R

Access the slides, code, and other materials at the course website at https://datavizm20.classes.andrewheiss.com/content/09-content/
  

https://youtu.be/lV_Qak75E6I
PMAP 8921 • (9) Annotations: (3) Seeds

Discussion of what seeds are and why they're important when dealing with pseudorandom numbers in R and ggplot2

Access the slides, code, and other materials at the course website at https://datavizm20.classes.andrewheiss.com/content/09-content/


https://youtu.be/gMSMsOy7KF0
PMAP 8921 • (9) Annotations: Example

Demonstration of how to create add text, shape, and arrow annotations to a ggplot, as well as enhance minor graphical elements and use ggtext to style text

Download the data and see the polished code at https://datavizm20.classes.andrewheiss.com/example/09-example/

```
