---
title: "Truth, beauty, and data revisited"
date: "2022-07-25"
date_end: "2022-07-30"
---

```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))
source(here::here("R", "youtube-playlist.R"))

playlist_id <- "PLS6tnpTr39sFYqZMtanRbtB1dKRBN2Rzt"

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "TDP_G1RiUmQ", "Introduction", "1", TRUE,
  "UyVJlbZjwhY", "Telling stories with data", "telling-stories", FALSE,
  "KHfF8zdxWzM", "Curiosity", "curiosity", FALSE
)
```

## Readings

- {{< fa book >}} Chapter 12 in *The Truthful Art* [@Cairo:2016]
- {{< fa book >}} [Chapter 26](https://clauswilke.com/dataviz/telling-a-story.html) in *Fundamentals of Data Visualization* [@Wilke:2018]
- {{< fa file-pdf >}} Martin Krzywinski and Alberto Cairo, ["Storytelling"](https://www.nature.com/articles/nmeth.2571.pdf) [@KrzywinskiCairo:2013]
- {{< fa brands youtube >}} Ben Wellington, ["Making data mean more through storytelling"](https://www.youtube.com/watch?v=6xsvGYIxJok)
- {{< fa brands vimeo >}} Jonathan Schwabish, ["Better Data Communication,"](https://vimeo.com/230757062) National Bureau of Economic Research
- {{< fa arrow-up-right-from-square >}} [Why People Make Bad Charts (and What to Do When it Happens)](https://flowingdata.com/2018/06/28/why-people-make-bad-charts-and-what-to-do-when-it-happens/)

### Questions to reflect on

*(Remember, you don't need to answer all of these—or even any of them! These are just here to help guide your thinking.)*

- Why are stories so powerful?
- How are stories related to truth?
- Is it ethical to emphasize certain aspects of the facts in data more than others? How do you decide which facts to use to convince audiences?
- When you're telling a story about data, you're inherently manipulating audience emotions. Is that okay?


## Slides

The slides for today's lesson are available online as an HTML file. Use the buttons below to open the slides either as an interactive website or as a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/15-slides")
slide_tabs(slide_details, "/slides/15-slides.html")
```

:::{.callout-tip}
**Fun fact**: If you type <kbd>?</kbd> (or <kbd>shift</kbd> + <kbd>/</kbd>) while going through the slides, you can see a list of special slide-specific commands.
:::


## Videos

```{r show-youtube-list, echo=FALSE, results="asis"}
youtube_list(slide_details, playlist_id)
```


```{cat include=FALSE, eval=FALSE}
https://www.youtube.com/playlist?list=PLS6tnpTr39sFYqZMtanRbtB1dKRBN2Rzt

https://youtu.be/TDP_G1RiUmQ
PMAP 8921 • (15) Truth, beauty, and data revisited: (0) Introduction

Introduction to the final class session

Access the slides, code, and other materials at the course website at https://datavizm20.classes.andrewheiss.com/content/15-content/


https://youtu.be/UyVJlbZjwhY
PMAP 8921 • (15) Truth, beauty, and data revisited: (1) Telling stories with data

Discussion of why it's important to tell stories with data and how to avoid major pitfalls related to manipulation, misinterpretation, ethos, and equity

Access the slides, code, and other materials at the course website at https://datavizm20.classes.andrewheiss.com/content/15-content/

  
https://youtu.be/KHfF8zdxWzM
PMAP 8921 • (15) Truth, beauty, and data revisited: (2) Curiosity

Discussion of how to continue to learn R by embracing your curiosity and working in public

Access the slides, code, and other materials at the course website at https://datavizm20.classes.andrewheiss.com/content/15-content/
```
